### 2020-07-14 transient.tex
Katy comment edits
final comment edits

This chapter discusses the transient multiphysics simulation results of the
\gls{MSFR} from Moltres for four accident scenarios. These scenarios, adapted
from Fiorina et al.'s work \cite{fiorina_modelling_2014}, include unprotected
instances of reactivity insertion, loss of heat sink, loss of flow, and pump
overspeed accidents. The term ``unprotected'' means no external interventions
occur in these scenarios. As such, these simulations
give an insight on the \gls{MSFR}'s passive safety capabilities in the absence
of any active safety system. This work used the steady-state configuration
presented in the previous chapter as the initial conditions for the transient
simulations discussed in this chapter. Specifically, all steady-state spatial
values for neutron flux, delayed neutron precursor concentration, temperature,
velocity, and pressure were imported as the initial state of the transient
scenarios.

As noted by Fiorina et al. \cite{fiorina_modelling_2014}, explicit decay heat
modeling has a negligible effect in reactivity-, and pump-initiated
transients. Furthermore, only their Polimi model had decay
heat modeling capabilities. Therefore, they presented results from their
Polimi and TUDelft model for the four accident scenarios without decay heat
modeling and enabled decay heat modeling for the loss of heat sink transient
scenario only. This work also ran all transient simulations without the decay
heat model for a fair comparison. The only exception is loss of heat sink
scenario in which two separate simulations with and without the decay heat
model were run. More generally, this work imposed various other simplifying
assumptions in our transient models to match their implementations as closely
as possible, within Moltres' capabilities. The details of the setup for each
transient simulation are in their respective sections.

\section{Unprotected Reactivity Insertion}

Reactivity insertion accidents are a type of nuclear accident caused by
unintended positive
reactivity insertions. The excess reactivity causes the power output and
temperatures in nuclear reactors to rise to potentially dangerous levels. In
\glspl{MSR}, a positive reactivity insertion could occur when the online
refueling system injects excess fissile material into the core. Excessively
high neutron fluences and temperatures negatively impact reactor structural
integrity and increase the risk of containment breach.

This work modeled two unprotected step-wise reactivity insertion scenarios in
Moltres by swapping out the original set of group constant data with two new,
separate sets of data from Serpent corresponding to 50 pcm and 200 pcm
reactivity insertions, respectively. The reactivity of the Serpent \gls{MSFR}
models was increased by increasing the $^{233}$U-to-$^{232}$Th ratio in
the fuel salt.

The focus of this transient study is the neutronic and thermal-hydraulic
behavior of the reactor core. Thus, this work assumes that the heat exchanger
and the associated power generation equipment (generator turbines, heat sinks,
etc.) can withstand all variations in the power output during the transients.

Figures \ref{fig:50pcmheat} and \ref{fig:50pcmtemp} show the power output and
average core temperature increase following the 50 pcm step-wise reactivity
insertion in the Moltres, Polimi, and TUDelft models. The initial prompt
response to the reactivity insertion raises power to 4 GW by $t=0.001$ s.
Figure \ref{fig:50pcmjump} shows the rise in power output specifically during
the prompt response in a separate plot.
Power continues to rise at a slower rate up to 4.63 GW at around $t=0.005$ s,
at which point the negative reactivity from the Doppler effect and salt
expansion becomes greater
than the initial +50 pcm insertion. Power continues to fall as the average
core temperature rises. The slight change in slope occurs at $t=0.3$ s. The
elapsed time approximately corresponds to the average half-life of the two
shortest-lived delayed neutron precursor (DNP) groups ($t_{1/2}=0.195$ s and
$0.424$ s); the decay of the surplus precursors produced in the initial phase
negates a fraction of the negative reactivity from the elevated core
temperature. By $t=3$-$4$ s, most of the heated salt and \glspl{DNP} from the
initial phase will have circulated around the primary loop and returned to the
core. The heated salt causes a small, noticeable dip in power before the power
output stabilises. The core tends to a new equilibrium average temperature
approximately 7.5 K higher than the initial average temperature.

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.85\textwidth]{50pcm-heat}
    \caption{Power output following
    a 50 pcm step-wise reactivity insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:50pcmheat}
%
    \centering
    \includegraphics[width=.85\textwidth]{50pcm-temp}
    \caption{Average core temperature increase following
    a 50 pcm step-wise reactivity insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:50pcmtemp}
\end{figure}

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.7\textwidth]{50pcm-jump}
    \caption{Power output during the prompt response following
    a 50 pcm step-wise reactivity insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:50pcmjump}
\end{figure}

@@

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.85\textwidth]{200pcm-heat}
    \caption{Power output following
    a 200 pcm step-wise reactivity insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:200pcmheat}
%
    \centering
    \includegraphics[width=.85\textwidth]{200pcm-temp}
    \caption{Average core temperature increase following
    a 200 pcm step-wise reactivity insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:200pcmtemp}
\end{figure}

The results for the 200 pcm reactivity insertion scenario show similar trends
to the 50 pcm case. The greater reactivity insertion elicits a stronger
prompt response in the power output which peaks at 92.1 GW. The average core
temperature increases much more rapidly and subsequently triggers a sharper
drop in power output. This results in the clearer distinction in the rate of
core temperature increase before and after $t=0.01$ s. In this transient, we
also observe greater deviation between Moltres and the other models arising
from the differences in the temperature reactivity coefficients. Overall,
Moltres' results show good agreement with the Polimi and TUDelft results.
The differences arise mainly due to the differences in the temperature
reactivity coefficients.

@@

An unprotected loss of heat sink accident can occur when the pumps in the
intermediate loop fail. The heat exchangers would then lose most of their
cooling capabilities. This work followed Fiorina et al.'s approach in assuming
that the cooling from the heat exchangers decreases exponentially with a time
constant of 1 s and all other parameters held constant
\cite{fiorina_modelling_2014}. As mentioned in the Chapter \ref{chap:ss}, we
will present two sets of results for this transient: 1) without decay heat
modeling, and 2) with decay heat modeling.

@@

Decay heat from fission products poses a great safety risk in a loss of
heat sink accident. Section \ref{sec:wodecayheat} showed that prompt fission
power output quickly falls as core temperatures rise. However, decay power
output is independent of the instantaneous neutron flux. Figure
\ref{fig:moltresdecaypower} shows that the decay power output remains
relatively high during a short-term transient. Decay heat becomes the dominant
heat source from $t=34$ s and falls at a much slower rate than prompt heat.
Figure \ref{fig:moltresdecaytemp} highlights the
greater core temperature increase attributed to decay heat as compared with
the results without decay heat. By $t=120$ s, the model with decay heat
records an average core temperature increase that is 45 K higher than the
model without decay heat. The average core temperature reaches approximately
1220 K and would continue to rise further. This places undue thermal stress
and accelerate corrosion rates in the Hastelloy structural material. In the
absence of an auxiliary heat removal system in the primary loop, reactor
operators would have to rely on the freeze plug to drain the core into a drain
tank with emergency cooling systems to keep the salt cool.

@@

A loss of forced flow transient can occur in the event of a station
blackout; the pumps would cease operating due to the loss of AC electrical
power. Natural circulation resulting from temperature-dependent density
changes becomes the dominant driving force for salt flow in the primary loop.
Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
approximation for buoyancy-driven flow in their models, but this approach was
not possible in Moltres because the primary loop is partitioned into two
separate geometries and used Dirichlet boundary
conditions at the inlet to drive flow. Fiorina et al.'s Polimi and TUDelft
models featured complete exponential coast-downs of the pumps with a time
constant of 5 s. The resulting flow rate $\dot{m}$ from natural circulation
was approximately 18 times smaller than the initial $\dot{m}$. Figure
\ref{fig:flowrate} shows that the actual $\dot{m}$ decreased with a time
constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, this work imposed
a similar exponential decay term with a time constant of 8 s on the inflow
Dirichlet boundary condition:
%
\begin{align}
    \text{Flow rate, } v = 0.25862 + (4.5-0.25862) e^{-t/8} 
    \label{eq:flowrate}
\end{align}

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.7\textwidth]{lof-flow-rate}
    \caption{The change in flow rate in the Polimi and TUDelft models and the
    imposed flow rate in Moltres.}
    \label{fig:flowrate}
\end{figure}

The reduced $\dot{m}$ also decreases the heat transfer rate between the
primary and intermediate loop through the heat exchanger as the heat transfer
coefficient $h$ is dependent on the $\dot{m}$. This step was problematic
because
the pointwise heat exchanger implementation in Moltres performs differently
compared with the heat exchangers that take up ``36\% of the out-of-core
part'' in the Polimi and TUDelft models \cite{fiorina_modelling_2014}. Most of
the cooling happens in the top half of the heat exchanger where the
temperature differential between the primary and intermediate loops is the
largest. In the Polimi and TUDelft models, the overall $h$ is the
``harmonic mean of the heat transfer coefficients on each side of the heat
exchanger''. For this loss of flow transient, the authors intended to focus on
the primary loop and assumed that only the pumps in the primary loop failed.
In addition to this, the authors applied the Dittus-Boelter correlation
\cite{dittus_heat_1930} for the relationship between the primary side $h$ and
$\dot{m}$. The Dittus-Boelter correlation for fluids being cooled is:
%
\begin{align}
    Nu &= 0.023 Re^{0.8} Pr^{0.3} \label{eq:db} \\
    \intertext{where}
    Nu &= \text{ Nusselt number,} \nonumber \\
    Re &= \text{ Reynolds number,} \nonumber \\
    Pr &= \text{ Prandtl number.} \nonumber
\end{align}
%
The only direct relation to $\dot{m}$ in the Dittus-Boelter correlation is
through the $Re$ term, which is directly proportional to flow velocity $v$.
This gives the following relation between $h$ and$v$:
%
\begin{align}
    h \propto v^{0.8} \label{eq:hv}
\end{align}
%
However, this relation provided very different results in the unprotected loss
of flow and pump overspeed transients compared with the Polimi and TUDelft
models; it overpredicted the equilibrium power output in the
unprotected loss of flow transient and underpredicted the same parameter in
the unprotected pump overspeed transient. Upon further investigation, the
present author found that raising the power of $v$ from 0.8 to 1.1 brought the
average core temperatures closer to the results from the other models in both
transients. Therefore, this work adopted the raised power in this study.

Another issue pertained to the turbulent viscosity $\mu_t$ as a function of
$v$. Using a simple approximation of $\mu_t$ being directly
proportional to $v$, the results differed significantly compared with the
Polimi and TUDelft models. This is likely due to buoyancy-driven flow
contributing to turbulence; the turbulent energy $k$ equation in COMSOL's
$k$-$\epsilon$ model has an explicit source term from buoyancy effects
\cite{comsol_ab_comsol_2018}. Another point to note is the
Reynolds number remains constant if $\mu$ and $v$ decrease in tandem. This
preserves the existence of the recirculation zone in the core and it is at
odds with the results from the Polimi and TUDelft models, which show that the
recirculation zones disappear during the loss of flow transient. We
circumvented this issue by letting fixed fractions of the initial $\mu_{t,0}$
be conserved regardless of the final flow velocity, according to the following
equation:
%
\begin{align}
    \mu_t &= \mu_c + (\mu_{t,0} - \mu_c) e^{-t/8} \label{lofmu} \\
    \intertext{where}
    \mu_c &= \text{ conserved fraction of $\mu_{t,0}$ [Pa$\cdot$s].} \nonumber
\end{align}
%
This measure allowed for laminar flow to develop in the core and yielded
results with closer resemblance to those from the Polimi and TUDelft models.

Figures \ref{fig:lofheat} and \ref{fig:loftemp} show the power output and
average core temperature increase during the unprotected loss of flow
transient in the Moltres, Polimi, and TUDelft models without decay heat
modeling. The three sets of results from Moltres correspond to $\mu_c =
\frac{1}{4} \mu_{t,0}, \frac{1}{2} \mu_{t,0}, \text{and } \frac{3}{4}
\mu_{t,0}$. Moltres
performs poorer in this transient relative to the two previous transients.
Although Moltres shows the same decreasing trend in power output, it failed to
capture the exact individual features in the reactor response. In the Polimi 
and TUDelft models, Fiorina et al.
stated that after around $t=15$ s, the ``flow pattern changed in the core and
the recirculation zones started to disappear''. A sudden drop in the average
core temperature results as the pocket of hot salt leaves the
core. In Moltres, the wider peak in the average core temperature
indicates that there was a more gradual change in the flow pattern.

Figure \ref{fig:lofflowtemp} shows the flow patterns and temperature
distribution in the core at $t=300$ s in all three models. Figures
\ref{fig:lofheat}, \ref{fig:loftemp}, and \ref{fig:lofflowtemp} combined
highlight the difference between laminar flow in the Moltres model and
buoyancy-driven flow in the other models, and its impact on the reactor
response. They show that low-speed laminar flow is a poor substitute for
buoyancy-driven flow in the context of the MSFR. It is particularly evident
in the transition from high-speed turbulent flow to low-speed viscous
flow as Moltres mispredicted the intermediate stages. The
simplifying assumption for the uniform, time-dependent $\mu_t$ is also flawed
in a safety analysis code.

@@

Pump overspeed refers to a sustained
increase in pump speed in the primary coolant loop. The increased flow rate
$\dot{m}$ impacts reactor performance in several ways.
It affects the neutronics by reducing the in-core $\beta$ as more of the
shorter-lived precursors will tend to flow out of the core before decaying.
This net loss of neutrons reduces the reactivity in the core, thereby causing
core temperatures to fall to counteract this change through temperature
reactivity feedback. The increased $\dot{m}$ also enhances the heat transfer
coefficient on the primary loop side of the heat exchanger and enables the
reactor to operate at a higher power output. At the same time, the improved
mixing flattens the temperature distribution in the core.

This workfollowed Fiorina et al.'s implementation
\cite{fiorina_modelling_2014} by
ramping up the inlet velocity, $u$, by 50\% from the nominal value, $u_0$,
according to the following formula:
%
\begin{align}
    u(t) &= u_0 [1 + 0.5 (1 - e^{-t / \tau})] \label{eq:overspeed}
    \intertext{where}
    \tau &= 5 \text{ s.} \nonumber
\end{align}
%
For this transient, this work assumed that $\mu_t$ was directly proportional
to $v$ because the buoyancy effects are negligible and the recirculation zones 
persist throughout the entire duration. Figures \ref{fig:poheat} and
\ref{fig:potemp} show the power output and average core temperature increase
during the unprotected pump overspeed transient in the Moltres, Polimi, and
TUDelft models. Figure \ref{fig:poshort} shows the same results for the first
20 seconds of the transient. At the start of the transient, the rising flow
rate cools the core and causes power output to rise sharply. Although the
average core temperature has a strictly decreasing trend, the temperature at
the center of the core briefly rises due ot the sharp increase in power
output. Since this is the region where most of the fissions take place, the
Doppler effect and salt expansion causes the power output to stall and dip
briefly before rising again at $t=2.5$ s. The reactor tends to a new
equilibrium power output and average core temperature. The temperature
distribution in the core is more evenly distributed because the turbulent
thermal conductivity $k_t$ is directly proportional to $mu_t$.

@@

The results show excellent agreement with the Polimi and TUDelft models. The
average core temperature increase in particular reproduces the Polimi and   
TUDelft results very well and the curve falls between the other two curves.
The power output is higher because the Moltres \gls{MSFR} model has a
stronger negative temperature coefficient than the other two models.

>>>>

This chapter discusses Moltres transient multiphysics simulations of the
\gls{MSFR} in four accident scenarios. These scenarios, adapted from the
\gls{MSFR} transient analyses with the Polimi and TUDelft models
\cite{fiorina_modelling_2014}, include unprotected reactivity
insertion, loss of heat sink, loss of flow, and pump overspeed accidents. The
term ``unprotected'' signify accident scenarios without treactor SCRAM. As
such, these simulations provide an insight on the
\gls{MSFR}'s passive safety capabilities in the absence
of active safety systems. This work used the steady-state configuration
presented in the previous chapter as the initial conditions for the transient
simulations discussed in this chapter. Specifically, all steady-state spatial
values for neutron flux, delayed neutron precursor concentration, temperature,
velocity, and pressure were imported as the initial state of the transient
scenarios.

As noted by Fiorina et al. \cite{fiorina_modelling_2014}, explicit decay heat
modeling has a negligible effect in reactivity-, and pump-initiated
transients. Furthermore, only their Polimi model had this capability.
Therefore, they featured decay heat modeling for the loss of heat sink
accident scenario only. The current work also ran all transient simulations
without the decay heat model for a fair comparison. The only exception is the
loss of heat sink scenario in which two separate simulations with and without
the decay heat model were run. This work imposed simplifying assumptions in
our transient models to match the implementations in
\cite{fiorina_modelling_2014} as closely as possible, within Moltres'
capabilities. The details of the setup for each
transient simulation appear in the following sections.

\section{Unprotected Reactivity Insertion}
In reactivity insertion accidents, excess reactivity would drive an increase
in flux, power, and temperature. In
\glspl{MSR}, a positive reactivity insertion could occur if the online
refueling system injected excess fissile material into the core. Excessively
high neutron fluences and temperatures could negatively impact reactor
structural integrity and increase the risk of containment breach.

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.7\textwidth]{insertion}
    \caption{Step-wise 50 pcm and 200 pcm reactivity insertions used to
    initiate the accident transients.}
    \label{fig:insertion}
\end{figure}

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.7\textwidth]{50pcm-jump}
    \caption{Power output during the prompt response following a 50 pcm
    step-wise unprotected reactivity insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:50pcmjump}
\end{figure}

This work modeled two unprotected step-wise reactivity insertion scenarios in
Moltres by swapping out the original set of group constant data with two new,
separate sets of data from Serpent corresponding to 50 pcm and 200 pcm
reactivity insertions, respectively. The reactivity of the Serpent \gls{MSFR}
models was increased by increasing the $^{233}$U-to-$^{232}$Th ratio in
the fuel salt. Figure \ref{fig:insertion} shows the step-wise reactivity
insertions used to model the accident scenarios.

The neutronic and thermal-hydraulic behaviors of the reactor core are the
focus of this transient study. Thus, this work assumes that the heat exchanger
and the associated power generation equipment (generator turbines, heat sinks,
etc.) can withstand all variations in the power output during the transients.

Figure \ref{fig:50pcmjump} shows the rise in power output during
the initial prompt response. The prompt
response raises the power output to 4 GW by $t=0.001$ s.
Figures \ref{fig:50pcmheat} and \ref{fig:50pcmtemp} show the power output and
average core temperature increase beyond $t=0.001$ s following the 50 pcm
step-wise reactivity insertion in the Moltres, Polimi, and TUDelft models.
Power continues to rise at a slower rate up to 4.63 GW at around $t=0.005$ s,
at which point the negative reactivity from the Doppler effect and salt
expansion becomes greater
than the initial +50 pcm insertion. Power continues to fall as the average
core temperature rises. A slight change in slope occurs at $t=0.3$ s. The
elapsed time approximately corresponds to the average half-life of the two
shortest-lived delayed neutron precursor (DNP) groups ($t_{1/2}=0.195$ s and
$0.424$ s); the decay of the surplus precursors produced in the initial phase
negates a fraction of the negative reactivity from the elevated core
temperature. By $t=3$ s, most of the heated salt and \glspl{DNP} from the
initial phase will have circulated around the primary loop and returned to the
core. The heated salt causes a small, noticeable dip in power before the power
output stabilizes.

The average core temperature rises steadily from the start of the transient
until $t=3$ s when the heated salt from the initial phase returns to the core.
This event is characterized by the small peak in the average core temperature
at $t=3$ s. The subsequent drop in power output halts the temperature
increase and the core tends to a new equilibrium average temperature
approximately 7.5 K higher than the initial average temperature.

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.85\textwidth]{50pcm-heat}
    \caption{Power output following a 50 pcm step-wise unprotected reactivity
    insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:50pcmheat}
%
    \centering
    \includegraphics[width=.85\textwidth]{50pcm-temp}
    \caption{Average core temperature increase following a 50 pcm step-wise
    unprotected reactivity insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:50pcmtemp}
\end{figure}

@@

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.85\textwidth]{200pcm-heat}
    \caption{Power output following a 200 pcm step-wise unprotected reactivity
    insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:200pcmheat}
%
    \centering
    \includegraphics[width=.85\textwidth]{200pcm-temp}
    \caption{Average core temperature increase following a 200 pcm step-wise
    unprotected reactivity insertion in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:200pcmtemp}
\end{figure}

The results for the 200 pcm reactivity insertion scenario show similar trends
to the 50 pcm case. The greater reactivity insertion elicits a stronger
prompt response in the power output which peaks at 92.1 GW. The average core
temperature increases much more rapidly and subsequently triggers a sharper
drop in power output. This more clearly distinguishes the rate of
core temperature increase before and after $t=0.01$ s. The difference in
$\alpha_T$ causes greater deviations in the results between Moltres and the
other models. Overall,
Moltres' results show good agreement with the Polimi and TUDelft results.
The differences arise mainly due to the differences in the temperature
reactivity coefficients.

@@

An unprotected loss of heat sink accident can occur when the pumps in the
intermediate loop fail and reactor is not SCRAMed. The heat exchangers would
then lose most of their
cooling capabilities. This work followed Fiorina et al.'s approach in assuming
that the cooling from the heat exchangers decreases exponentially with a time
constant of 1 s and all other parameters held constant
\cite{fiorina_modelling_2014}. As mentioned in the Chapter \ref{chap:ss}, we
will present two sets of results for this transient: 1) without decay heat
modeling, and 2) with decay heat modeling.

@@

Decay heat from fission products poses a great safety risk in an unprotected
loss of
heat sink accident. Section \ref{sec:wodecayheat} showed that prompt fission
power output quickly falls as core temperatures rise. However, decay power
output is independent of the instantaneous neutron flux. Figure
\ref{fig:moltresdecaypower} shows that the decay power output remains
relatively high during a short-term transient. Decay heat becomes the dominant
heat source from $t=34$ s and falls at a much slower rate than prompt heat.
Figure \ref{fig:moltresdecaytemp} highlights the greater core temperature
increase arising from decay heat as compared with
the results without decay heat. By $t=120$ s, the model with decay heat
records an average core temperature increase that is 45 K higher than the
model without decay heat. The absolute core temperature reaches approximately
1220 K and would continue to rise further. This places undue thermal stress
and accelerates salt-induced corrosion in the Hastelloy structural material.
In the
absence of an auxiliary heat removal system in the primary loop, reactor
operators would have to rely on the freeze plug to drain the core into a drain
tank with emergency cooling systems to keep the salt cool.

@@

@@

Pump overspeed refers to a sustained
increase in pump speed in the primary coolant loop. The increased flow rate
$\dot{m}$ impacts reactor performance in several ways.
It affects the neutronics by reducing the in-core $\beta$ as more of the
shorter-lived precursors will tend to flow out of the core before decaying.
This net loss of neutrons reduces the reactivity in the core, thereby causing
core temperatures to fall to counteract this change through temperature
reactivity feedback. The increased $\dot{V}$ also enhances the heat transfer
coefficient on the primary loop side of the heat exchanger and enables the
reactor to operate at a higher power output. At the same time, the improved
mixing flattens the temperature distribution in the core.

This work followed Fiorina et al.'s implementation
\cite{fiorina_modelling_2014} by
ramping up the inlet velocity, $u$, by 50\% from the nominal value, $u_0$,
according to the following formula:
%
\begin{align}
    u(t) &= u_0 [1 + 0.5 (1 - e^{-t / \tau})] \label{eq:overspeed}
    \intertext{where}
    \tau &= 5 \text{ s.} \nonumber
\end{align}
%
For this transient, this work assumed that $\mu_t$ was directly proportional
to $v$ because the buoyancy effects are assumed to be negligible with forced
flow and the recirculation zones 
persist throughout the entire duration.

Setting the exact dependence between the heat transfer coefficient $h$ and
$\dot{V}$ was problematic because
the pointwise heat exchanger implementation in Moltres performs differently
compared with the heat exchangers of finite volume in the Polimi and TUDelft
models \cite{fiorina_modelling_2014}. In a heat exchanger of finite volume,
most of the cooling occurs in the first half of the heat exchanger where the
temperature differential between the primary and intermediate loops is the
largest. In the Polimi and TUDelft models, the overall $h$ is a
harmonic mean of the $h_i$ on each side of the heat exchanger, given as:
%
\begin{align}
    h &= \frac{2}{\frac{1}{h_1} + \frac{1}{h_2}}, \\
    \intertext{where}
    h &= \text{overall heat transfer coefficient
    [W$\cdot$K$^{-1}$],} \nonumber \\
    h_1 &= \text{heat transfer coefficient on the primary loop side
    [W$\cdot$K$^{-1}$],} \nonumber \\
    h_2 &= \text{heat transfer coefficient on the intermediate loop side
    [W$\cdot$K$^{-1}$].} \nonumber
\end{align}
%
In addition to this, Fiorina et al. applied the Dittus-Boelter correlation
\cite{dittus_heat_1930} for the relationship between $h_1$ and
$\dot{V}$. The Dittus-Boelter correlation for fluids being cooled is:
%
\begin{align}
    Nu &= 0.023 Re^{0.8} Pr^{0.3}, \label{eq:db} \\
    \intertext{where}
    Nu &= \text{Nusselt number,} \nonumber \\
    &\text{\ \ \ \ \ the ratio of convective to conductive heat
    transfer at a boundary in a fluid,} \nonumber \\
    Re &= \text{Reynolds number,} \nonumber \\
    &\text{\ \ \ \ \ the ratio of inertial forces to viscous
    forces within a fluid,} \nonumber \\
    Pr &= \text{Prandtl number,} \nonumber \\
    &\text{\ \ \ \ \ the ratio of momentum diffusivity to thermal
    diffusivity.} \nonumber
\end{align}
%
The only direct relation to $\dot{V}$ in the Dittus-Boelter correlation is
through the Reynolds number, $Re$, which is directly proportional to
flow velocity $v$. This gives the following relation between $h$ and $v$:
%
\begin{align}
    h &\propto v^{0.8} \label{eq:hv} \\
    h &= \text{heat transfer coefficient [W$\cdot$K$^{-1}$],} \nonumber \\
    v &= \text{flow velocity [m$\cdot$s$^{-1}$].} \nonumber
\end{align}
%
However, this relation provided very different results in the unprotected pump
overspeed and loss of flow transients compared with the Polimi and TUDelft
models. This approach underpredicted the equilibrium power output in the
unprotected pump overspeed transient and overpredicted the same parameter in
the unprotected loss of flow transient. Upon further investigation, I
found that raising the power of $v$ from 0.8 to 1.1 brought the
average core temperatures closer to the results from the other models in both
transients. Therefore, this work adopted the raised power in this thesis.

Figures \ref{fig:poheat} and
\ref{fig:potemp} show the power output and average core temperature increase
during the unprotected pump overspeed transient in the Moltres, Polimi, and
TUDelft models. Figure \ref{fig:poshort} shows the same results for the first
20 seconds of the transient. At the start of the transient, the rising flow
rate cools the core and causes power output to rise sharply. Although the
average core temperature has a strictly decreasing trend, the temperature at
the center of the core briefly rises due to the sharp increase in power
output. Since this is the region where most of the fissions take place, the
Doppler effect and salt expansion causes the power output to stall and dip
briefly before rising again at $t=2.5$ s. The reactor tends to a new
equilibrium power output and average core temperature. The temperature
distribution in the core is more evenly distributed because the turbulent
thermal conductivity $k_t$ is directly proportional to $mu_t$.

@@

In both sets of results, Moltres reproduced all of the transient features
found in the Polimi and TUDelft models. The average core temperature profile
falls between the Polimi and TUDelft results, while the power output is
approximately 0.1 GW higher because the Moltres \gls{MSFR} model has a
greater $\alpha_T$ than the other two models.

A loss of forced flow transient can occur in the event of a station blackout
without reactor SCRAM; the pumps would cease operating due to the loss of AC
electrical
power. Natural circulation resulting from temperature-dependent density
changes becomes the dominant driving force for salt flow in the primary loop.
Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
approximation for buoyancy-driven flow in the Polimi and TUDelft models, but
this approach was not possible in Moltres because the primary loop is
partitioned into two separate geometries and used Dirichlet boundary
conditions at the inlet to drive flow. The Polimi and TUDelft
models featured complete exponential coast-downs of the pumps with a time
constant of 5 s. The final flow rate $\dot{V}_f$ from natural circulation
was approximately 18 times smaller than the initial $\dot{V}_0$. Figure
\ref{fig:flowrate} shows that the actual $\dot{V}$ decreased with a time
constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, this work imposed
a similar exponential decay term with a time constant of 8 s on the inflow
Dirichlet boundary condition:
%
\begin{align}
    \dot{V} &= \dot{V}_f + (\dot{V}_0-\dot{V}_f) e^{-t/8},
    \label{eq:flowrate} \\
    \intertext{where}
    \dot{V}_f &= \text{final volumetric flow rate} = 0.25862
    \text{ m$^3\cdot$s$^{-1}$,} \nonumber \\
    \dot{V}_0 &= \text{initial volumetric flow rate} = 4.5
    \text{ m$^3\cdot$s$^{-1}$.} \nonumber
\end{align}

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.7\textwidth]{lof-flow-rate}
    \caption{The change in flow rate in the Polimi and TUDelft models and the
    imposed flow rate in Moltres.}
    \label{fig:flowrate}
\end{figure}
%
The reduced $\dot{V}$ also decreases the heat transfer rate between the
primary and intermediate loop through the heat exchanger as the heat transfer
coefficient $h$ is dependent on the $\dot{V}$. For this loss of flow
transient, Fiorina et al. \cite{fiorina_modelling_2014} intended to focus
on the primary loop and assumed that only the pumps in the primary loop
failed.

An issue arose pertaining to the turbulent viscosity $\mu_t$ as a function of
$v$. When using a simple approximation of $\mu_t$ being directly proportional
to $v$, the results from Moltres differed significantly compared with the
Polimi and TUDelft models. The difference is due to buoyancy-driven flow
contributing to turbulence; the turbulent energy $k$ equation in COMSOL's
$k$-$\epsilon$ model has an explicit source term from buoyancy effects
\cite{comsol_ab_comsol_2018}. Another point to note is the
Reynolds number remains constant if $\mu$ and $v$ decrease in tandem. This
preserves the existence of the recirculation zone in the core and it is at
odds with the results from the Polimi and TUDelft models, which show that the
recirculation zones disappear during the loss of flow transient. The current
work circumvented this issue by letting fixed fractions of the initial
$\mu_{t,0}$ be conserved regardless of the final flow velocity, according to
the following equation:
%
\begin{align}
    \mu_t &= \mu_c + (\mu_{t,0} - \mu_c) e^{-t/8} \label{lofmu} \\
    \intertext{where}
    \mu_c &= \text{ conserved fraction of $\mu_{t,0}$ [Pa$\cdot$s].} \nonumber
\end{align}
%
This measure allowed for laminar flow to develop in the core and yielded
results showing better qualitative agreement with those from the Polimi and
TUDelft models. The subsequent paragraphs discuss these results.

Figures \ref{fig:lofheat} and \ref{fig:loftemp} show the power output and
average core temperature increase during the unprotected loss of flow
transient in the Moltres, Polimi, and TUDelft models without decay heat
modeling. The three sets of results from Moltres correspond to $\mu_c =
\frac{1}{4} \mu_{t,0}, \frac{1}{2} \mu_{t,0}, \text{and } \frac{3}{4}
\mu_{t,0}$.
Although Moltres shows the same decreasing trend in power output, it failed to
capture the exact individual features in the reactor response. In the Polimi 
and TUDelft models, Fiorina et al. \cite{fiorina_modelling_2014}
stated that after around $t=15$ s, the ``flow pattern changed in the core and
the recirculation zones started to disappear''. The pocket of hot salt leaves
the core and causes a sudden drop in the average core temperature. In Moltres,
the wider peak in the average core temperature
indicates that there was a more gradual change in the flow pattern.

Figure \ref{fig:lofflowtemp} shows the flow patterns and temperature
distribution in the core at $t=300$ s in all three models. Figures
\ref{fig:lofheat}, \ref{fig:loftemp}, and \ref{fig:lofflowtemp} combined
highlight the difference between laminar flow in the Moltres model and
buoyancy-driven flow in the other models, and its impact on the reactor
response. They show that low-speed laminar flow is a poor substitute for
buoyancy-driven flow in the context of the MSFR. It is particularly evident
in the transition from high-speed turbulent flow to low-speed viscous
flow as Moltres mispredicted the intermediate stages. As for $\mu_t$, the need
to fine-tune this parameter confirms that the uniform, function-based $\mu_t$
approach is inappropriate for safety analysis in a loss of flow accident.
